---
title: "SSB Descriptive Data"
author: "Meghan Visnick Davis"
date: "2025-05-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      echo = FALSE)
```

# Cohort

Of the N = 1,090 observations in the provided dataset, n = 988 had some available exposure data, n = 467 had some available outcome data, and n = 442 had both. The following descriptive stats apply to the cohort of n = 442.

```{r}
# load libraries 
library(tidyverse)
library(mosaic)
library(lattice)
library(descr)
library(flextable)

# read in data
ssb_cohort <- read.csv('ssb_cohort.csv')
```

# Exposure Data

Values for "_ECHO" variables are missing for most of the cohort. These frequencies are still presented alongside their counterpart variables. 

SUGARFRUIT_DRINK_ECHO & HIGHCAFFEIN_SODA_ECHO values are missing for the entire original dataset of n = 1090 - possible data extraction error?

## Sugar-sweetened soda 

"During the past month, how often did you drink regular, carbonated pop, soda, or soft drinks that contain sugar?  Do not include diet soda.  READ IF NECESSARY:   Do not include diet or sugar-free fruit drinks.  Do not include juices or tea in cans.  Do not include diet mineral water or diet flavored waters."

0 = Never ; 
1 = 1-3 times last month ; 
2 = 1-2 times per week ; 
3 = 3-4 times per week ; 
4 = 5-6 times per week ; 
5 = 1 time per day ; 
6 = 2 times per day ; 
7 = 3 times per day ; 
8 = 4 times per day ; 
9 = 5 or more times per day ; 
98 = Don't Know ; 
99 = Refuse 

```{r}
# convert to factor variable
ssb_cohort$SUGARSODA <- as.factor(ssb_cohort$SUGARSODA)

# frequency tables
freq(ssb_cohort$SUGARSODA,plot = FALSE)
freq(ssb_cohort$SUGARSODA_ECHO,plot = FALSE)

# bar chart
plt <- ggplot(ssb_cohort, aes(x = SUGARSODA)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Sugar-sweetened soda",
       y = "Frequency")
plt
```

## Fruit juice

"During the past month, how often did you drink 100% fruit juice, such as orange, mango, apple, and grape juices?  Do not include fruit drinks.   READ IF NECESSARY:   Include only 100% pure juices. Do not include fruit drinks with added sugar, like Kool-Aid®, Hi-C®, lemonade, cranberry cocktail, Gatorade®, Tampico®, and Sunny Delight®."

0 = Never ; 
1 = 1-3 times last month ; 
2 = 1-2 times per week ; 
3 = 3-4 times per week ; 
4 = 5-6 times per week ; 
5 = 1 time per day ; 
6 = 2 times per day ; 
7 = 3 times per day ; 
8 = 4 times per day ; 
9 = 5 or more times per day ; 
98 = Don't Know ; 
99 = Refuse 

```{r}
# convert to factor variable
ssb_cohort$FRUITJUICE <- as.factor(ssb_cohort$FRUITJUICE)

# frequency tables
freq(ssb_cohort$FRUITJUICE,
     plot = FALSE)
freq(ssb_cohort$FRUITJUICE_ECHO,
     plot = FALSE)

# bar chart
plt <- ggplot(ssb_cohort, aes(x = FRUITJUICE)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "100% Fruit juice",
       y = "Frequency")
plt
```

## Fruit-flavored drinks with added sugar

"During the past month, how often did you drink fruit-flavored drinks with sugar (such as Kool-Aid®, Hi-C®, lemonade, or cranberry cocktail)?  Do not include diet drinks.  READ IF NECESSARY:   Include Gatorade® and other sports drinks with added sugar.   Include Tampico®, Sunny Delight® and Twister®.   Do not include 100% fruit juices or soda.  Do not include yogurt drinks or carbonated water."

0 = Never ; 
1 = 1-3 times last month ; 
2 = 1-2 times per week ; 
3 = 3-4 times per week ; 
4 = 5-6 times per week ; 
5 = 1 time per day ; 
6 = 2 times per day ; 
7 = 3 times per day ; 
8 = 4 times per day ; 
9 = 5 or more times per day ; 
98 = Don't Know ; 
99 = Refuse 

```{r}
# convert to factor variable
ssb_cohort$SUGARFRUIT_DRINK <- as.factor(ssb_cohort$SUGARFRUIT_DRINK)

# frequency table
freq(ssb_cohort$SUGARFRUIT_DRINK,
     plot = FALSE)

# bar chart
plt <- ggplot(ssb_cohort, aes(x = SUGARFRUIT_DRINK)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Fruit-flavored drinks with added sugar",
       y = "Frequency")
plt
```


## Highly-caffeinated sodas

"During the past month, how often did you drink Jolt®, Surge®, Mountain Dew®, Red Bull® and other highly caffeinated sodas?"

0 = Never ; 
1 = less than once per month ; 
2 = 1-3 times per month ; 
3 = 1 time per week ; 
4 = 2-4 times per week ; 
5 = 5-6 times per week ; 
6 = 1 time per day ; 
7 = 2-3 times per day ; 
8 = 4-5 times per day ; 
9 = 6 or more times per day ; 
98 = Don't Know ; 
99 = Refuse 

```{r}
# convert to factor variable
ssb_cohort$HIGHCAFFEIN_SODA <- as.factor(ssb_cohort$HIGHCAFFEIN_SODA)

# frequency table
freq(ssb_cohort$HIGHCAFFEIN_SODA,
     plot = FALSE)

# bar chart
plt <- ggplot(ssb_cohort, aes(x = HIGHCAFFEIN_SODA)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Highly-caffeinated sodas",
       y = "Frequency")
plt
```

## Sweetened fruit drinks & Sugar sweetened coffee or tea

During the past month, how often did you drink sweetened fruit drinks, sports or energy drinks, such as Kool-Aid, lemonade, Hi-C, cranberry drink, Gatorade, Red Bull, or Vitamin Water? Include fruit juices you made at home and added sugar to. Do not include diet drinks or artificially sweetened drinks.

1 = Never ; 2 = 1 time last month ; 3 = 2-3 times last month ; 4 = 1 time per week ; 5 = 2 times per week ; 6 = 3-4 times per week ; 7 = 5-6 times per day ; 8 = 1 time per day; 9 = 2-3 times per day ; 10 = 4-5 tiems per day ; 11 = 6 or more times per day ; 98 = Don't Know ; 99 = Refuse 

During the past month, how often did you drink coffee or tea that had sugar or honey added to it? Include coffee and tea you sweetened yourself and presweetened tea and coffee drinks such as Arizona Iced Tea or Frappuccino. Do not include artificially sweetened coffee or diet tea.

1 = Never ;  2 = 1 time last month ; 3 = 2-3 times last month ; 4 = 1 time per week ; 5 = 2 times per week ; 6 = 3-4 times per week ; 7 = 5-6 times per week ; 8 = 1 time per day ; 8 = 2-3 times per day ; 10 = 4-5 times per day ; 11 = 6 or more times per day ; 98 = Don't Know ; 99 = Refuse

Data almost entirely missing for these two. 

```{r}
# SWEETDRINK_ECHO

## convert to factor variable
ssb_cohort$SWEETDRINK_ECHO <- as.factor(ssb_cohort$SWEETDRINK_ECHO)

## frequency table
freq(ssb_cohort$SWEETDRINK_ECHO,
     plot = FALSE)

# SUGAR_HONEY_ECHO

## convert to factor variable
ssb_cohort$SUGAR_HONEY_ECHO <- as.factor(ssb_cohort$SUGAR_HONEY_ECHO)

## frequency table
freq(ssb_cohort$SUGAR_HONEY_ECHO,
     plot = FALSE)
```

# Outcome Data

Availability of CBQ data was largely what determined the size of this cohort. I followed the instrument scoring guidance to create score variables, which are the variables presented here. Per the guidance, scores were only included if less than or equal to 20% of items were missing or marked as not applicable. 

ASQ-3 data was missing for about half the cohort (n = 225, n = 223, n = 223 for communication, problem solving, and personal-social, respectively). I only included the adjusted total scores here, after confirming that all adjusted totals were valid scores per the scoring guidance (with no more than 2 questions missing). 

I also noticed there are two other ASQ-3 elements - gross motor and fine motor. I'm assuming these weren't of interest?

## CBQ - Total score

```{r}
# cbq summary stats
result <- favstats(ssb_cohort$cbq_total)

result[] <- lapply(result, function(x) {
  if (is.numeric(x)) round(x, 2) else x
})

flextable(result)

# histogram
histogram(ssb_cohort$cbq_total,
          breaks = 30, 
          main = "Distribution of CBQ - Total score", 
          ylab = 'percent', 
          type = 'percent')


```

## CBQ - Negative affect

```{r}
# cbq summary stats
result <- favstats(ssb_cohort$cbq_na)

result[] <- lapply(result, function(x) {
  if (is.numeric(x)) round(x, 2) else x
})

flextable(result)

# histogram
histogram(ssb_cohort$cbq_na, 
          breaks = 30,
          main = "Distribution of CBQ - Negative affect score", 
          ylab = 'percent', 
          type = 'percent')
```


## CBQ - Surgency

```{r}
# cbq summary stats
result <- favstats(ssb_cohort$cbq_surg)

result[] <- lapply(result, function(x) {
  if (is.numeric(x)) round(x, 2) else x
})

flextable(result)

# histogram
histogram(ssb_cohort$cbq_surg,
          breaks = 30,
          main = "Distribution of CBQ - Surgency", 
          ylab = 'percent', 
          type = 'percent')
```


## CBQ - Effort control 

```{r}
# cbq summary stats
result <- favstats(ssb_cohort$cbq_ec)

result[] <- lapply(result, function(x) {
  if (is.numeric(x)) round(x, 2) else x
})

flextable(result)

# histogram
histogram(ssb_cohort$cbq_ec, 
          breaks = 30,
          main = "Distribution of CBQ - Effort control", 
          ylab = 'percent', 
          type = 'percent')
```

## ASQ-3 - Communication

```{r}
# asq-3 summary stats
result <- favstats(ssb_cohort$asq_36_total_communication)

result[] <- lapply(result, function(x) {
  if (is.numeric(x)) round(x, 2) else x
})

flextable(result)

# histogram
histogram(ssb_cohort$asq_36_total_communication, 
          breaks = 20,
          main = "Distribution of ASQ-3 - Communication", 
          ylab = 'percent', 
          type = 'percent')
```

## ASQ-3 - Problem solving

```{r}
# asq-3 summary stats
result <- favstats(ssb_cohort$asq_36_total_problemsolving)

result[] <- lapply(result, function(x) {
  if (is.numeric(x)) round(x, 2) else x
})

flextable(result)

# histogram
histogram(ssb_cohort$asq_36_total_problemsolving, 
          breaks = 20,
          main = "Distribution of ASQ-3 - Problem solving", 
          ylab = 'percent', 
          type = 'percent')
```

## ASQ-3 Personal-social

```{r}
# asq-3 summary stats
result <- favstats(ssb_cohort$asq_36_total_personal_social)

result[] <- lapply(result, function(x) {
  if (is.numeric(x)) round(x, 2) else x
})

flextable(result)

# histogram
histogram(ssb_cohort$asq_36_total_personal_social, 
          breaks = 20,
          main = "Distribution of ASQ-3 - Personal-social", 
          ylab = 'percent', 
          type = 'percent')
```

# Potential confounders

## Maternal age at childbirth

Depending on the rationale for maternal age as a confounder, I'd consider treating this as a categorical variable and recoding with age groups, or at least plot age against both exposure and outcome variables to explore further - rationale being that the relationship between age and outcomes might be nonlinear. This could of course be true with any continuous variable, but particularly with maternal age during pregnancy we tend to see morbidity at both extremes of age. 

```{r}
# maternal age at birth summary stats
result <- favstats(ssb_cohort$maternal_age_birth)

result[] <- lapply(result, function(x) {
  if (is.numeric(x)) round(x, 2) else x
})

flextable(result)

# histogram
histogram(ssb_cohort$maternal_age_birth, 
          breaks = 20,
          main = "Maternal age at birth (years)", 
          ylab = 'percent', 
          type = 'percent')
```


## Mother's race

Lots of small numbers, would need to combine categories here - probably just White, Black, Other. 

```{r}
# frequency table
freq(ssb_cohort$Mother_Race,
     plot = FALSE)

# bar chart
plt <- ggplot(ssb_cohort, aes(x = Mother_Race)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Race of mother",
       x = "Race",
       y = "Frequency")
plt + coord_flip()

```

## Mother's ethnicity

I'd probably drop this as a variable, due to not enough observations in the hispanic/latina group. 

```{r}
# frequency table
freq(ssb_cohort$Mother_Ethnicity,
     plot = FALSE)

# bar chart
plt <- ggplot(ssb_cohort, aes(x = Mother_Ethnicity)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Ethnicity of mother",
       x = "Ethnicity",
       y = "Frequency")
plt + coord_flip()
```

## Height, Weight, Body mass index (BMI)

I used height and weight variables to derive pre-pregnancy body mass index (BMI). Cut points for the categorical BMI variable are based on WHO/NIH clinical BMI classifications. 

```{r}
# create new height variable combining HEIGHT_FEET & HEIGHT_INCH
ssb_cohort$height_inches <- ssb_cohort$HEIGHT_FEET*12 + ssb_cohort$HEIGHT_INCH

# recode extreme weight outliers to missing
ssb_cohort$PREWT_LBS[ssb_cohort$PREWT_LBS == 998] <- NA

# create new variable for BMI
ssb_cohort$prepreg_bmi <- (ssb_cohort$PREWT_LBS/(ssb_cohort$height_inches^2))*703

# categorical bmi 
ssb_cohort$bmi_cat <- cut(ssb_cohort$prepreg_bmi,
                          breaks = c(-Inf, 18.5, 25, 30, Inf),
                          labels = c("Underweight", "Normal", "Overweight", "Obese"),
                          right = FALSE)  # Set to TRUE if upper bound is inclusive
```

### Height (inches)

```{r}
# height
result <- favstats(ssb_cohort$height_inches)

result[] <- lapply(result, function(x) {
  if (is.numeric(x)) round(x, 2) else x
})

flextable(result)

# histogram
histogram(ssb_cohort$height_inches,
          breaks = 20, 
          main = "Distribution of height (inches)", 
          ylab = 'percent', 
          type = 'percent')
```

### Weight (lbs)

There were 8 data entries of "998" lbs. The spread of weights all made sense otherwise. I excluded these from the summary stats. 

```{r}
# pre-preg weight
result <- favstats(ssb_cohort$PREWT_LBS)

result[] <- lapply(result, function(x) {
  if (is.numeric(x)) round(x, 2) else x
})

flextable(result)

# histogram
histogram(ssb_cohort$PREWT_LBS,
          breaks = 20, 
          main = "Distribution of pre-pregnancy weight (lbs)", 
          ylab = 'percent', 
          type = 'percent')
```

### Pre-pregnancy BMI

```{r}
# bmi
result <- favstats(ssb_cohort$prepreg_bmi)

result[] <- lapply(result, function(x) {
  if (is.numeric(x)) round(x, 2) else x
})

flextable(result)

# histogram
histogram(ssb_cohort$prepreg_bmi,
          breaks = 25, 
          main = "Distribution of BMI", 
          ylab = 'percent', 
          type = 'percent')

# bmi category
## frequency table
freq(ssb_cohort$bmi_cat,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = bmi_cat)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Pre-pregnancy BMI",
       y = "Frequency")
plt + coord_flip()
```


## Marital status

Are you currently married, living with a partner, divorced, separated, widowed, or have you never been married?
1 = Married ; 2 = Living with a partner ; 3 = Divorced ; 4 = Separated ; 5 = Widowed ; 6 = Never married ; 98 = Don't know ; 99 = Refuse

Lots of small numbers, would need to combine categories. 

```{r}
# convert to character variable
ssb_cohort$MARIT_STATUS <- as.factor(ssb_cohort$MARIT_STATUS)

# frequency table
freq(ssb_cohort$MARIT_STATUS,
     plot = FALSE)

# bar chart
plt <- ggplot(ssb_cohort, aes(x = MARIT_STATUS)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Marital status",
       y = "Frequency")
plt 

```

## Parity

Number of pregnancies carried to >20 weeks. Almost all are missing. 

```{r}
# convert to character variable
ssb_cohort$mmra_b3c <- as.factor(ssb_cohort$mmra_b3c)

# frequency table
freq(ssb_cohort$mmra_b3c,
     plot = FALSE)
```

## Child Race

Like maternal race, I'd collapse this to White, Black, or Other. 

```{r}
# frequency table
freq(ssb_cohort$Child_Race,
     plot = FALSE)

# bar chart
plt <- ggplot(ssb_cohort, aes(x = Child_Race)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Race of child",
       x = "Race",
       y = "Frequency")
plt + coord_flip()
```

## Child Ethnicity

Like maternal ethnicity, the numbers are  small in the hispanic/latino(a) group.

```{r}
# frequency table
freq(ssb_cohort$Child_Ethnicity,
     plot = FALSE)

# bar chart
plt <- ggplot(ssb_cohort, aes(x = Child_Ethnicity)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Ethnicity of child",
       x = "Ethnicity",
       y = "Frequency")
plt + coord_flip()
```

## Child - sex

Two variables:

dem_c_2: What is the sex of the child at the time of birth?
1=Male |  2= Female | 3=Ambiguous |  7= Prefer not to answer, 8=Don't know

cbra_c2: Sex of infant
1= Male | 2= Female | 3= Ambiguous | -8= Not documented

I was surprised these numbers weren't more even; I included the frequencies on the whole dataset here for comparison. 

```{r}
# explore sex variables
freq(ssb_cohort$dem_c_2, 
     plot = FALSE)          
freq(ssb_cohort$cbmra_c2,
     plot = FALSE)          # most are missing

# bar chart
plt <- ggplot(ssb_cohort, aes(x = dem_c_2)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Sex of child",
       y = "Frequency")
plt

# explore sex variables on entire dataset
ssb <- read.csv('C:/Users/Meghan/Documents/Professional/SSB Paper/Initial files from Kelly/Hirko08022024_Dataset 2.csv')
freq(ssb$dem_c_2,
     plot = FALSE)
freq(ssb$cbmra_c2,
     plot = FALSE)
```

## Child - birthweight

I ran summary stats for all birthweight variables and didn't go through the trouble to label appropriately here - essentially all are missing. 

```{r}
favstats(ssb_cohort$cbmra_c7a_grams)
favstats(ssb_cohort$cbmra_c7c_lbs)
favstats(ssb_cohort$cbmra_c7c_oz)
```

## Child - gestational age at birth (weeks)

```{r}
# asq-3 summary stats
result <- favstats(ssb_cohort$GA_Birth)

result[] <- lapply(result, function(x) {
  if (is.numeric(x)) round(x, 2) else x
})

flextable(result)

# histogram
histogram(ssb_cohort$GA_Birth, 
          breaks = 20,
          main = "Distribution of gestational age at birth (weeks)", 
          ylab = 'percent', 
          type = 'percent')
```

## Maternal educational level

What is the highest grade or level of school you have completed or the highest degree you have received?

1 = None ; 2 = 8th grade or less ; 3 = Some high school, no diploma ; 4 = High school graduate, diploma, or GED ; 5 = Some college credit, no degree ; 6 = Trade/Technical/Vocational training ; 7 = Associate degree ; 8 = Bachelor's degree ; 9 = Master's degree ; 10 = Doctorate or professional degree ; 98 = don't know ; 99 = refuse 

```{r}
# convert to factor variable
ssb_cohort$EDUC_LVL <- as.factor(ssb_cohort$EDUC_LVL)

# frequencies
freq(ssb_cohort$EDUC_LVL, 
     plot = FALSE)           


# bar chart
plt <- ggplot(ssb_cohort, aes(x = EDUC_LVL)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Maternal educational level",
       y = "Frequency")
plt
```

## Most recent income

```{r}
# frequencies
freq(ssb_cohort$final_income, 
     plot = FALSE)           

# bar chart
plt <- ggplot(ssb_cohort, aes(x = final_income)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Most recent income",
       y = "Frequency")
plt + coord_flip()
```

## Age at weaning

There were no data available for this cohort (Variable: MTHR_BRST_STOPAGE_ECHO). 

## Substance use in pregnancy

### Alcohol

During this pregnancy, how many alcoholic drinks have you had in an average week?

1 = 14 drinks or more a week ; 2 = 7 to 13 drinks a week ; 3 = 4 to 6 drinks a week ; 4 = 1 to 3 drinks a week ; 5 = Less than 1 drink a week ; 6 = I didn't drink ; 98 = Don't know ; Refuse

```{r}
## convert to categorical 
ssb_cohort$ALCOH_DRINKPREG <- as.factor(ssb_cohort$ALCOH_DRINKPREG)

## frequencies
freq(ssb_cohort$ALCOH_DRINKPREG,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = ALCOH_DRINKPREG)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Alcohol use",
       y = "Frequency")
plt + coord_flip()
```

### Hookah

During this pregnancy, have you used a hookah?

1 = Yes ; 5 = No ; 98 = Don't know ; 99 = Refuse

```{r}
# convert to categorical 
ssb_cohort$HOOKAH_PREG <- as.factor(ssb_cohort$HOOKAH_PREG)

# frequencies
freq(ssb_cohort$HOOKAH_PREG,
     plot = FALSE)

# bar chart
plt <- ggplot(ssb_cohort, aes(x = HOOKAH_PREG)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Hookah use",
       y = "Frequency")
plt + coord_flip()
```

### E-cigarettes

During this pregnancy, have you used e-cigarettes or other electronic nicotine products?

1 = Yes ; 5 = No ; 98 = Don't know ; 99 = Refuse

```{r}
# convert to categorical 
ssb_cohort$ECIGARETTE_PREG <- as.factor(ssb_cohort$ECIGARETTE_PREG)

# frequencies
freq(ssb_cohort$ECIGARETTE_PREG,
     plot = FALSE)

# bar chart
plt <- ggplot(ssb_cohort, aes(x = ECIGARETTE_PREG)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "E-cigarette use",
       y = "Frequency")
plt + coord_flip()
```

### Cigarettes

Which of the following best describes your cigarette smoking?

1 = I smoke regularly now; about the same as before finding out I was pregnant ; 2 = I smoke regularly now, but I've cut down since I found out I was pregnant = ; 3 = I smoke every once in a while ; 4 = I have quite smoking since finding out I was pregnant ; 5 = I wasn't smoking around the time I found out I was pregnant, and I don't currently smoke cigarettes ; 98 = Don't know ; 99 = Refuse 

```{r}
# convert to categorical 
ssb_cohort$CIGARETTE_SMOKING <- as.factor(ssb_cohort$CIGARETTE_SMOKING)

# frequencies
freq(ssb_cohort$CIGARETTE_SMOKING,
     plot = FALSE)

# bar chart
plt <- ggplot(ssb_cohort, aes(x = CIGARETTE_SMOKING)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Cigarette smoking",
       y = "Frequency")
plt + coord_flip()
```

### Amphetamines

Have you used Amphetamine type stimulants (e.g., “uppers”, Dexedrine, speed, Benzadrine, ecstasy) at all during this pregnancy?

1 = Yes ; 5 = No ; 98 = Don't know ; 99 = Refuse

(Same question and answer coding for both - looks like one is asked at the first prenatal visit and the other is asked at the second)

```{r}
# prenatal 1

## convert to categorical 
ssb_cohort$AMPHE_PREG <- as.factor(ssb_cohort$AMPHE_PREG)

## frequencies
freq(ssb_cohort$AMPHE_PREG,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = AMPHE_PREG)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Amphetamine use - Prenatal 1",
       y = "Frequency")
plt + coord_flip()

# prenatal 2

## convert to categorical 
ssb_cohort$AMPHE_PREG2 <- as.factor(ssb_cohort$AMPHE_PREG2)

## frequencies
freq(ssb_cohort$AMPHE_PREG2,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = AMPHE_PREG2)) + 
  geom_bar(fill = "#F8766D") +               
  labs(title = "Amphetamine use - Prenatal 2",
       y = "Frequency")
plt + coord_flip()
```

### Steroids

Have you used steroids (e.g., dexamethasone, hydrocortisone, prednisone) at all during this pregnancy?

1 = Yes ; 5 = No ; 98 = Don't know ; 99 = Refuse

```{r}
# prenatal 1

## convert to categorical 
ssb_cohort$STEROID_PREG <- as.factor(ssb_cohort$STEROID_PREG)

## frequencies
freq(ssb_cohort$STEROID_PREG,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = STEROID_PREG)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Steroid use - Prenatal 1",
       y = "Frequency")
plt + coord_flip()

# prenatal 2

## convert to categorical 
ssb_cohort$STEROID_PREG2 <- as.factor(ssb_cohort$STEROID_PREG2)

## frequencies
freq(ssb_cohort$STEROID_PREG2,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = STEROID_PREG2)) + 
  geom_bar(fill = "#F8766D") +               
  labs(title = "Steroid use - Prenatal 2",
       y = "Frequency")
plt + coord_flip()
```

### Tranquilizers

Have you used tranquilizers (e.g., Xanax, Ativan, Valium, Buspar,) at all during this pregnancy?

1 = Yes ; 5 = No ; 98 = Don't know ; 99 = Refuse

```{r}
# convert to categorical 
ssb_cohort$TRANQUI_PREG <- as.factor(ssb_cohort$TRANQUI_PREG)

# frequencies
freq(ssb_cohort$TRANQUI_PREG,
     plot = FALSE)

# bar chart
plt <- ggplot(ssb_cohort, aes(x = TRANQUI_PREG)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Tranquilizer use",
       y = "Frequency")
plt + coord_flip()
```

### Marijuana

Have you used marijuana (pot) or cannabis, including medically prescribed cannabis at all during this pregnancy?

1 = Yes ; 5 = No ; 98 = Don't know ; 99 = Refuse

```{r}
# prenatal 1

## convert to categorical 
ssb_cohort$MARIJ_PREG <- as.factor(ssb_cohort$MARIJ_PREG)

## frequencies
freq(ssb_cohort$MARIJ_PREG,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = MARIJ_PREG)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Marijuana use - Prenatal 1",
       y = "Frequency")
plt + coord_flip()

# prenatal 2

## convert to categorical 
ssb_cohort$MARIJ_PREG2 <- as.factor(ssb_cohort$MARIJ_PREG2)

## frequencies
freq(ssb_cohort$MARIJ_PREG2,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = MARIJ_PREG2)) + 
  geom_bar(fill = "#F8766D") +               
  labs(title = "Marijuana use - Prenatal 2",
       y = "Frequency")
plt + coord_flip()
```

### Cocaine

Have you used cocaine (e.g., powdered cocaine; crack; coke) at all during this pregnancy?

1 = Yes ; 5 = No ; 98 = Don't know ; 99 = Refuse

```{r}
# prenatal 1

## convert to categorical 
ssb_cohort$COCAINE_PREG <- as.factor(ssb_cohort$COCAINE_PREG)

## frequencies
freq(ssb_cohort$COCAINE_PREG,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = COCAINE_PREG)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Cocaine use - Prenatal 1",
       y = "Frequency")
plt + coord_flip()

# prenatal 2

## convert to categorical 
ssb_cohort$COCAINE_PREG2 <- as.factor(ssb_cohort$COCAINE_PREG2)

## frequencies
freq(ssb_cohort$COCAINE_PREG2,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = COCAINE_PREG2)) + 
  geom_bar(fill = "#F8766D") +               
  labs(title = "Cocaine use - Prenatal 2",
       y = "Frequency")
plt + coord_flip()
```

### Methamphetamine

Have you used methamphetamine (e.g., crystal, ice) at all during this pregnancy?

1 = Yes ; 5 = No ; 98 = Don't know ; 99 = Refuse

```{r}
# prenatal 1

## convert to categorical 
ssb_cohort$METHAM_PREG <- as.factor(ssb_cohort$METHAM_PREG)

## frequencies
freq(ssb_cohort$METHAM_PREG,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = METHAM_PREG)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Methampetamine use - Prenatal 1",
       y = "Frequency")
plt + coord_flip()

# prenatal 2

## convert to categorical 
ssb_cohort$METHAM_PREG2 <- as.factor(ssb_cohort$METHAM_PREG2)

## frequencies
freq(ssb_cohort$METHAM_PREG2,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = METHAM_PREG2)) + 
  geom_bar(fill = "#F8766D") +               
  labs(title = "Methamphetmine use - Prenatal 2",
       y = "Frequency")
plt + coord_flip()
```

### Prescription Opioids

Have you used prescription opioids (e.g. codeine, MS Contin, oxycontin, hydrocodone or Vicodin, oxycodone or Percocet, darvocet, demerol, morphine, fentanyl) at all during this pregnancy?

1 = Yes ; 5 = No ; 98 = Don't know ; 99 = Refuse

```{r}
# prenatal 1

## convert to categorical 
ssb_cohort$OPIOID_PREG <- as.factor(ssb_cohort$OPIOID_PREG)

## frequencies
freq(ssb_cohort$OPIOID_PREG,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = OPIOID_PREG)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Prescription Opioid use - Prenatal 1",
       y = "Frequency")
plt + coord_flip()

# prenatal 2

## convert to categorical 
ssb_cohort$OPIOID_PREG2 <- as.factor(ssb_cohort$OPIOID_PREG2)

## frequencies
freq(ssb_cohort$OPIOID_PREG2,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = OPIOID_PREG2)) + 
  geom_bar(fill = "#F8766D") +               
  labs(title = "Prescription Opioid use - Prenatal 2",
       y = "Frequency")
plt + coord_flip()
```

### Non-prescription Opioids

Have you used non-prescription opioids (heroin)  at all during this pregnancy?

1 = Yes ; 5 = No ; 98 = Don't know ; 99 = Refuse

*Side note: I think it's really interesting the way this and the previous questions are worded - what about people who have used opioid pills (not heroin) but without an Rx? 

```{r}
# prenatal 1

## convert to categorical 
ssb_cohort$NOPRESCOPIOID_PREG <- as.factor(ssb_cohort$NOPRESCOPIOID_PREG)

## frequencies
freq(ssb_cohort$NOPRESCOPIOID_PREG,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = NOPRESCOPIOID_PREG)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Non-prescription Opioid use - Prenatal 1",
       y = "Frequency")
plt + coord_flip()

# prenatal 2

## convert to categorical 
ssb_cohort$NOPRESCOPIOID_PREG2 <- as.factor(ssb_cohort$NOPRESCOPIOID_PREG2)

## frequencies
freq(ssb_cohort$NOPRESCOPIOID_PREG2,
     plot = FALSE)

## bar chart
plt <- ggplot(ssb_cohort, aes(x = NOPRESCOPIOID_PREG2)) + 
  geom_bar(fill = "#F8766D") +               
  labs(title = "Non-prescription Opioid use - Prenatal 2",
       y = "Frequency")
plt + coord_flip()
```

### Opioid addiction treatments

Have you used opioid addiction treatments (e.g. methadone, buprenorphine) at all during this pregnancy?

1 = Yes ; 5 = No ; 98 = Don't know ; 99 = Refuse

```{r}
# convert to categorical 
ssb_cohort$OPIOIDTREAT_PREG2 <- as.factor(ssb_cohort$OPIOIDTREAT_PREG2)

# frequencies
freq(ssb_cohort$OPIOIDTREAT_PREG2,
     plot = FALSE)

# bar chart
plt <- ggplot(ssb_cohort, aes(x = OPIOIDTREAT_PREG2)) + 
  geom_bar(fill = "#00BA38") +               
  labs(title = "Use of Opioid addiction treatment",
       y = "Frequency")
plt + coord_flip()
```

```{r}
# write new dataset
write.csv(ssb_cohort,"ssb_cohort_2.csv") 

```

